---
title: "DATA 621 Final Project"
#author: "Marco Siqueira Campos"
date: "April 25th, 2018"
output: 
        pdf_document:
                latex_engine: xelatex
mainfont: Calibri Light 
fontsize: 12pt

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

####DATA 621  
####MSDS CUNY  

#### Final Project 
Wisconsin Diagnostic Breast Cancer (WDBC)

### Team: Christopher Estevez and Marco Siqueira Campos.   
 
 
### Project Description

PLEASE CHRISTOPHER HERE

Features are computed from a digitized image of a fine needle aspirate (FNA) of a breast mass. They describe characteristics of the cell nuclei present in the image. 

The mean, standard error, and "worst" or largest (mean of the three
largest values) of these features were computed for each image,
resulting in 30 features.  For instance, field 3 is Mean Radius, field
13 is Radius SE, field 23 is Worst Radius.

### 1. DATA EXPLORATION 

Variable description

1) ID number
2) diagnosis (M = malignant, B = benign), this we changed for (1 = malignant, 0 = benign).

Ten real-valued features are computed for each cell nucleus: 

a) radius (mean of distances from center to points on the perimeter)
b) texture (standard deviation of gray-scale values)
c) perimeter
d) area
e) smoothness (local variation in radius lengths)
f) compactness (perimeter^2 / area - 1.0)
g) concavity (severity of concave portions of the contour)
h) concave points (number of concave portions of the contour)
i) symmetry
j) fractal dimension ("coastline approximation" - 1)

For each variable we have 3 different statistics: Mean, standard error and worst case.
We have a total of 30 predictors.

The change `diagnosis` predictor from malignant for 1, benign for 0, help us to see relationship beetwen the response and the predictors, and was done in this phase.


```{r, echo=FALSE, warning=FALSE, message=FALSE}
#load libraries
library(xlsx)
library(psych)
library(GGally)
library(ggplot2)
library(caret)
library(dplyr)
library(pROC)
library(car)
library(leaps)
library(sjstats)
```

```{r, echo=FALSE, cache=TRUE}
#load dataset 
WDBCdata<-read.xlsx("data_with groups.xlsx", sheetName = "ics data")
```

```{r, echo=FALSE, cache=TRUE}
# First step the id will be removed to avoid any future trouble
# change M = malignant for 1 and B for B = benign for zero
WDBCdata$mycol<-NULL
WDBCdata$id<-NULL
WDBCdata$diagnosis<-ifelse(WDBCdata$diagnosis == "B", 0, 1)
```

Summary statistics

```{r, echo=FALSE}
#summary stats
WDBC_tbl<-describe(WDBCdata,IQR=T)[,c(1:5,8:10,11,12)]
knitr::kable(round(WDBC_tbl,2), caption = "Selected Stats")
rm(WDBC_tbl)
```

Histograms

```{r, echo=FALSE, fig.width=7, fig.height=8, message=F, warning=F}
library(Hmisc)
hist.data.frame(WDBCdata, n.unique=1, mtitl = "Breast Cancer Histogram")
```
\begin{center}
fig. 1
\end{center}


BOX-PLOT

```{r ,echo=FALSE, fig.align="center", fig.height=4, fig.width=6}
# boxplot
par(cex.axis=0.8) # is for x-axis
boxplot(WDBCdata[,c(2:15)], las=2, col="red", main="Breast Cancer Box-Plot" )
boxplot(WDBCdata[,c(16:31)], las=2, col="red", main="Breast Cancer Box-Plot" )

```
\begin{center}
fig. 2
\end{center}

BOX-PLOT

```{r ,echo=FALSE, fig.align="center", fig.height=4, fig.width=6}
# boxplot
par(cex.axis=0.8) # is for x-axis
boxplot(WDBCdata[,c(2:15)], las=2, col="green", main="Breast Cancer Box-Plot, y axis changed", ylim = c(0,150))
boxplot(WDBCdata[,c(16:31)], las=2, col="green", main="Breast Cancer Box-Plot, y axis changed", ylim = c(0,150))

```
\begin{center}
fig. 2
\end{center}


Correlation matrix

```{r ,echo=FALSE}
ggcorr(WDBCdata, nbreaks=8, palette='PRGn', label=TRUE, label_size=2, size = 1.8, label_color='black') + ggtitle("Breast Cancer Correlation Matrix") + theme(plot.title = element_text(hjust = 0.5, color = "grey15"))
```

Correlation list (only high correlation values, r>0.9)
```{r ,echo=FALSE}
z = cor(WDBCdata)

z = round(z,4)
z[abs(z)<0.9]=NA # remove low relationship
z[lower.tri(z,diag=TRUE)]=NA  #Prepare to drop duplicates and meaningless information
z=as.data.frame(as.table(z))  #Turn into a 3-column table
z=na.omit(z)  #Get rid of the junk we flagged above
z=z[order(-abs(z$Freq)),]    #Sort by highest correlation (whether +ve or -ve)
z
rm(z)
```

Summary of data explorations findings:

- Most histograms present very asymmetric behavior with similar to exponential distribution.
- Some predictor look like exponential distribuition as `radius_se`, `perimeter_se`, `area_se`, `concavity_se` and `fractal_dimension_se`.
- There is no true outliers, the outliers at box-plot is due the kind of distribution.
- There is no missing.
- we identified 21 pairs of highly correlated predictors, r> 0.9, this was due to the choice of predictors that are associated, measures things related: radius, perimeter and area.
- There are 14 predictors related with the response, `Diagnosis`, with r>=0.6, this is a good news.  

###2. DATA PREPARATION

The to solve the issues the follow activities will be done:

- Solve high correlations values, the follow predictors were removed: `area_mean`, `radius_mean`, `area_worst`, `compactness_mean`, `perimeter_worst` and `compactness_se`. -concavity_worst -fractal_dimension_worst.
- Verify the data behavior of power transformation for the follow predictors: `radius_se`, `perimeter_se`, `area_se`, `concavity_se` and `fractal_dimension_se`.

Power transformation check (HERE IS THE SAME, WE NEED TEST IS THE CHANGE IMPROVE OR NOT THE MODEL, HERE IS ONLY TO CHECK IF THE POWER CAN IMPROVE THE DISTRIBUTION)

```{r ,echo=FALSE,fig.align="center", fig.height=4, fig.width=6}

bc1<-BoxCoxTrans(WDBCdata$radius_se)
bc2<-BoxCoxTrans(WDBCdata$perimeter_se)
bc3<-BoxCoxTrans(WDBCdata$area_se )
#bc4<-BoxCoxTrans(WDBCdata$concavity_se)
bc5<-BoxCoxTrans(WDBCdata$fractal_dimension_se)

par(mfrow=c(1,2))
hist(WDBCdata$radius_se)
hist(WDBCdata$radius_se^bc1$lambda)
par(mfrow=c(1,2))
hist(WDBCdata$perimeter_se)
hist(WDBCdata$perimeter_se^bc2$lambda)
par(mfrow=c(1,2))
hist(WDBCdata$area_se)
hist(WDBCdata$area_se^bc3$lambda)
par(mfrow=c(1,2))
hist(WDBCdata$fractal_dimension_se)
hist(WDBCdata$fractal_dimension_se^bc5$lambda)
par(mfrow=c(1,1))

```


```{r, echo=FALSE}
# code to compute VIF to indentify predictors to drop

#vif_logit<-glm(WDBCdata$diagnosis ~. -diagnosis,family=binomial,data=WDBCdata)
#vif(vif_logit)
```



```{r, echo=FALSE, cache=TRUE}

WDBCdatafull<-WDBCdata #copy data set

WDBCdata<-subset(WDBCdata, select=-c(area_mean,radius_mean,area_worst,compactness_mean,perimeter_worst,compactness_se,concavity_worst,fractal_dimension_worst)) # remove predictors


# WDBCdatafull all predictors
# WDBCdata remove highly correlated predictors


set.seed(123)
indx<-createDataPartition(WDBCdata$diagnosis, p=0.7, list=FALSE)

train_x<-WDBCdata[indx,-1]
train_y<-WDBCdata$diagnosis[indx]

test_x<-WDBCdata[-indx,-1]
test_y<-WDBCdata$diagnosis[-indx]


# the subset bellow is with full predictors, cannot be used with logistic model, only can be used with model that accept highly correlated predictors.


trainfull_x<-WDBCdatafull[indx,-1]
trainfull_y<-WDBCdatafull$diagnosis[indx]

testfull_x<-WDBCdatafull[-indx,-1]
testfull_y<-WDBCdatafull$diagnosis[-indx]

```        






REFERENCES:  
[1] Dua, D. and Karra Taniskidou, E. (2017). UCI Machine Learning Repository [http://archive.ics.uci.edu/ml]. Irvine, CA: University of California, School of Information and Computer Science.    
[2] https://stackoverflow.com/questions/7074246/show-correlations-as-an-ordered-list-not-as-a-large-matrix

