---
title: "Predicting the breast cancer by characteristics of the cell nuclei present in the image"
author: "Marco Siqueira Campos,Shyam BV,Christopher Estevez,Ahmed Sajjad,Aryeh Sturm"
date: "`r format(Sys.Date(), format='%B %d, %Y')`"
output: 
  pdf_document:
    toc: yes
    toc_depth: 3
mainfont: Calibri Light 
fontsize: 12pt

---

```{r global_options, echo=TRUE, include=TRUE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```

#Abstract

 
#Introduction and Background

*MOTIVATION*  
*WHAT IS THE QUESTION?*

Features are computed from a digitized image of a fine needle aspirate (FNA) of a breast mass. They describe characteristics of the cell nuclei present in the image. 

The mean, standard error, and "worst" or largest (mean of the three
largest values) of these features were computed for each image,
resulting in 30 features.  For instance, field 3 is Mean Radius, field
13 is Radius SE, field 23 is Worst Radius.

#Methodology



##DATA EXPLORATION 

Variable description

1) ID number
2) diagnosis (M = malignant, B = benign), this we changed for (1 = malignant, 0 = benign).

Ten real-valued features are computed for each cell nucleus: 

a) radius (mean of distances from center to points on the perimeter)
b) texture (standard deviation of gray-scale values)
c) perimeter
d) area
e) smoothness (local variation in radius lengths)
f) compactness (perimeter^2 / area - 1.0)
g) concavity (severity of concave portions of the contour)
h) concave points (number of concave portions of the contour)
i) symmetry
j) fractal dimension ("coastline approximation" - 1)

For each variable we have 3 different statistics: Mean, standard error and worst case.
We have a total of 30 predictors.

The change `diagnosis` predictor from malignant for 1, benign for 0, help us to see relationship beetwen the response and the predictors, and was done in this phase.


```{r, message=TRUE, warning=TRUE}
#load libraries
library("xlsx")
library("psych")
library("GGally")
library("ggplot2")
library("caret")
library("dplyr")
library("pROC")
library("car")
library("leaps")
library("sjstats")
library("Hmisc")
library("PerformanceAnalytics")
library("factoextra")
library("BGLR")
library("keras")
library("dplyr")
#library("kableExtra")
#install.packages("Hmisc")
```


Lets load the dataset from the excel spreadsheet.

```{r, echo=FALSE, cache=TRUE}
#load dataset 
WDBCdata<- read.xlsx("data/data_with groups.xlsx", sheetName = "ics data")
```

### Data Analysis

Below is the structure and the summary of the dataset.


```{r}
str(WDBCdata)
summary(WDBCdata)
```

Below are the inferenace of the summary.

1. `diagonisis` is the target variable.
2. `myCol and id` are index. It is not required in predictors.
3. There is no `NA` in all the predictors. Imputation is not required.
4. Each variable is in different scale.
5. Each cell nuclei has different properties like `radius, texture, perimeter, area, smoothness, compactness, concavity, concave points, symmetry_mean, fractal dimenstion`. Dataset also has Mean, standard error and the worst measure of that particular cell.

As a next step we will remove the unwanted variables and analyze individual set of dimensions.

### Data Visualization


```{r, echo=FALSE, cache=TRUE}
# First step the id will be removed to avoid any future trouble
# change M = malignant for 1 and B for B = benign for zero
WDBCdata$mycol<-NULL
WDBCdata$id<-NULL
WDBCdata$diagnosis<-ifelse(WDBCdata$diagnosis == "B", 0, 1)
```



#### Histogram

As a first visualization, we will plot the hisogram of all the predictor variables.

```{r, echo=FALSE}
#summary stats
WDBC_tbl<-describe(WDBCdata,IQR=T)[,c(1:5,8:10,11,12)]
#knitr::kable(round(WDBC_tbl,2), caption = "Selected Stats")
rm(WDBC_tbl)

```

Histograms

```{r, echo=FALSE, fig.width=7, fig.height=8, message=F, warning=F}

hist.data.frame(WDBCdata, n.unique=1, mtitl = "Breast Cancer Histogram")


```
\begin{center}
fig. 1
\end{center}




#### BOX-PLOT


Lets split the predictors according to its category


```{r}
WDBCdata_mean = cbind(diagnosis=WDBCdata[,c(1)], WDBCdata[,c(2:11)])
WDBCdata_se = cbind(diagnosis=WDBCdata[,c(1)], WDBCdata[,c(12:21)])
WDBCdata_worst = cbind(diagnosis=WDBCdata[,c(1)], WDBCdata[,c(22:31)])

```


BOX-PLOT

```{r ,echo=FALSE, fig.align="center", fig.height=4, fig.width=6}
# boxplot
par(cex.axis=0.8) # is for x-axis
boxplot(WDBCdata_mean, las=2, col="green", main="Breast Cancer Box-Plot for Mean", ylim = c(0,150))
boxplot(WDBCdata_se, las=2, col="green", main="Breast Cancer Box-Plot for SE", ylim = c(0,150))
boxplot(WDBCdata_worst, las=2, col="green", main="Breast Cancer Box-Plot for Worst", ylim = c(0,150))

```
  
\begin{center}
fig. 3
\end{center}

Some inference from the chart

1. Nuclei mean of the perimeter and area is higher.
2. Standard Error of the Area is higher.
3. In worst nuclei scenario, area has extremly high values. 


#### Correlation matrix

Now lets plot the correlation plots to understand more about the correlated predictors.


```{r}
#Mean
chart.Correlation(WDBCdata_mean,histogram=TRUE,pch=19)

```

```{r}
# SE
chart.Correlation(WDBCdata_se,histogram=TRUE,pch=19)
```


```{r}
# Worst
chart.Correlation(WDBCdata_worst,histogram=TRUE,pch=19)
```

Below is the overall correlation matrix of all the predictors


```{r ,echo=FALSE}
ggcorr(WDBCdata, nbreaks=8, palette='PRGn', label=TRUE, label_size=2, size = 1.8, label_color='black') + ggtitle("Breast Cancer Correlation Matrix") + theme(plot.title = element_text(hjust = 0.5, color = "grey15"))
```

\begin{center}
fig. 4
\end{center}

Correlation list (only high correlation values, r>0.9)
```{r ,echo=FALSE}
z = cor(WDBCdata)

z = round(z,4)
z[abs(z)<0.9]=NA # remove low relationship
z[lower.tri(z,diag=TRUE)]=NA  #Prepare to drop duplicates and meaningless information
z=as.data.frame(as.table(z))  #Turn into a 3-column table
z=na.omit(z)  #Get rid of the junk we flagged above
z=z[order(-abs(z$Freq)),]    #Sort by highest correlation (whether +ve or -ve)
z
rm(z)


```

#### Diagnosis Plots

Lets deepdive into response variable and see its distribution.

```{r}
qplot(as.factor(WDBCdata$diagnosis))+geom_bar() + labs(x='Diagnosis', y ='Count')
```



Summary of data explorations findings:

- Most histograms present very asymmetric behavior with similar to exponential distribution.
- Some predictor look like exponential distribuition as `radius_se`, `perimeter_se`, `area_se`, `concavity_se` and `fractal_dimension_se`.
- There is no true outliers, the outliers at box-plot is due the kind of distribution.
- There is no missing.
- we identified 21 pairs of highly correlated predictors, r> 0.9, this was due to the choice of predictors that are associated, measures things related: radius, perimeter and area.
- There are 14 predictors related with the response, `Diagnosis`, with r>=0.6, this is a good news.  

## DATA PREPARATION

To solve the highly correlated variables, we will follow two types. 

1. Predictors transformation and remove correlated variables.
2. PCA transformed correlated variables.

### Predictors transformation and remove correlated variables

In this method following steps will be performed:

- Solve high correlations values, the follow predictors were removed: `area_mean`, `radius_mean`, `area_worst`, `compactness_mean`, `perimeter_worst`, `compactness_se`. `concavity_worst`and `fractal_dimension_worst`.
- Verify the data behavior of power transformation for the follow predictors: `radius_se`, `perimeter_se`, `area_se` and `fractal_dimension_se`.


There are some predictor variables which are skwed. We can perform power transformations on those variables.

```{r ,echo=FALSE,fig.align="center", fig.height=3, fig.width=5}

bc1<-BoxCoxTrans(WDBCdata$radius_se)
bc2<-BoxCoxTrans(WDBCdata$perimeter_se)
bc3<-BoxCoxTrans(WDBCdata$area_se )
bc5<-BoxCoxTrans(WDBCdata$fractal_dimension_se)

par(mfrow=c(1,2))
hist(WDBCdata$radius_se, main="Histogram radius_se")
hist(WDBCdata$radius_se^bc1$lambda, main="Histogram radius_se transf.")
par(mfrow=c(1,2))
hist(WDBCdata$perimeter_se, main="Histogram perimeter_se")
hist(WDBCdata$perimeter_se^bc2$lambda, main="Histogram perimeter_se transf.")
par(mfrow=c(1,2))
hist(WDBCdata$area_se, main="Histogram area_se")
hist(WDBCdata$area_se^bc3$lambda, main="Histogram area_se transf.")
par(mfrow=c(1,2))
hist(WDBCdata$fractal_dimension_se, main="Histogram dimension_se")
hist(WDBCdata$fractal_dimension_se^bc5$lambda, main="Histogram dimension_se transf.")
par(mfrow=c(1,1))

```
\begin{center}
fig. 3
\end{center}

As we saw in the chart, figure 3, the power transformation, has improved the distribution.


```{r, echo=FALSE}
# code to compute VIF to indentify predictors to drop

#vif_logit<-glm(WDBCdata$diagnosis ~. -diagnosis,family=binomial,data=WDBCdata)
#vif(vif_logit)
```

Now we will remove the highly correlated variables and then split the data in two data set, Train data set with 70% and Test data set with 30% of all data. 

```{r, echo=FALSE, cache=TRUE}

WDBCdatafull<-WDBCdata #copy data set

WDBCdata<-subset(WDBCdata, select=-c(area_mean,radius_mean,area_worst,compactness_mean,perimeter_worst,compactness_se,concavity_worst,fractal_dimension_worst)) # remove predictors


# WDBCdatafull all predictors
# WDBCdata remove highly correlated predictors


set.seed(123)
indx<-createDataPartition(WDBCdata$diagnosis, p=0.7, list=FALSE)

train_x<-WDBCdata[indx,-1]
train_y<-WDBCdata$diagnosis[indx] 

test_x<-WDBCdata[-indx,-1]
test_y<-WDBCdata$diagnosis[-indx]


# the subset bellow is with full predictors, cannot be used with logistic model, only can be used with model that accept highly correlated predictors.

trainfull_x<-WDBCdatafull[indx,-1]
trainfull_y<-WDBCdatafull$diagnosis[indx]

testfull_x<-WDBCdatafull[-indx,-1]
testfull_y<-WDBCdatafull$diagnosis[-indx]

```        


```{r, echo=FALSE, cache=TRUE}
# applying the power transformation for the test and train data set

train_x$radius_se<-train_x$radius_se^bc1$lambda
train_x$perimeter_se<-train_x$perimeter_se^bc2$lambda    
train_x$area_se<-train_x$area_se^bc3$lambda
train_x$fractal_dimension_se<-train_x$fractal_dimension_se^bc5$lambda

test_x$radius_se<-test_x$radius_se^bc1$lambda
test_x$perimeter_se<-test_x$perimeter_se^bc2$lambda    
test_x$area_se<-test_x$area_se^bc3$lambda
test_x$fractal_dimension_se<-test_x$fractal_dimension_se^bc5$lambda

``` 

### PCA transformed correlated variables

In this type, as the variables are highly correlated, we will transform the predictors using principal component analysis(PCA). PCA will provide the transformed variables.

```{r}

pca_wdbc <- prcomp(train_x[,2:ncol(train_x)],center = TRUE, scale=TRUE)

plot(pca_wdbc, type='l')

summary(pca_wdbc$x)
```

Above summary shows that 10 PCA's shows the 95% of variation. And 17 PCA's shows 99.1% of variation in the dataset. We will select 17 PCA's to show the variation in the data. It also has a clusters of around 7. Now lets see the clusters in detail.

```{r}


pca_wdbc_var <- get_pca_var(pca_wdbc)


res <- kmeans(pca_wdbc_var$coord,centers = 5, nstart=25)
grp <- as.factor(res$cluster)

fviz_pca_var(pca_wdbc, col.var=grp, palette='jco', legend.title='Cluster')

```





##MODELS 
###Logistic Regression

Now we will build different models and compare the results. 

As a initial set of models, we will try logistic regression with different variations in it.

#### **Logistic Regression Model(Logit) with all variables**

As a first model, we will build logistic regression model which has all the predictors with correlated variables..

```{r}
model_11_logit_full <-glm(trainfull_y ~ . ,family=binomial, trainfull_x)
summary(model_11_logit_full)
```

None of the predictors are signficant due to correlated variables. This model did not provide any results and did not converge.

#### Logistic Regression without correlated variables

```{r}
model_12_logit_corr <-glm( train_y~.,family=binomial,data=train_x)
summary(model_12_logit_corr)

```

This time the model did converge, however it seems many variables are not significant. We will perform backward stepwise selection of the model.

```{r}
model_13_logit_corr_final <- update(model_12_logit_corr, .~.-symmetry_se-concave.points_worst-texture_se-perimeter_se-radius_se-symmetry_mean-fractal_dimension_mean-concave.points_mean-texture_mean-compactness_worst-symmetry_worst-smoothness_mean-area_se-smoothness_se )
summary(model_13_logit_corr_final)
```

Now all the variables are statistically significant and AIC got reduced. 

Following predictors were chosen:
- `perimeter_mean` 
- `concavity_mean`
- `concavity_se`
- `concave.points_se`
- `fractal_dimension_se`
- `radius_worst` 
- `texture_worst`
- `smoothness_worst`

Now lets try out other models.

#### Logistic regression on PCA variables

We can also perform logistic regression on PCA transformed variables. However, the problem is the model is not interpretable. One of the biggest advantage of Logistic regression is interpretability. By using PCA variables, we loose that advantag. But lets see how the model reacts to the PCA variables.


```{r}

model_14_pca <-glm( train_y~.,family=binomial,data=pca_wdbc$x[,c(1:14)] %>% data.frame())
summary(model_14_pca)

```
PCA models are difficult to interpret. So this model does not provide any other information. So we will not proceed with this type of model.

#### Logistic Regression - Probit

Now lets perform logistic regression with probit link function. After backward stepwise eliminiation, below is the model we got.

```{r}
model_15_probit <-glm( train_y~.-perimeter_se-texture_worst-concave.points_worst-symmetry_se-fractal_dimension_mean-symmetry_mean-compactness_worst-concavity_se-concave.points_se-concavity_mean-smoothness_se-texture_se-symmetry_worst,family=binomial(link = 'probit'),data=train_x)
summary(model_15_probit)
```
Seems there are nine predictors which are significant. It has similar AIC score compared to logit model.

### Bayesian Binary regression

Bayesian Genralized linear regression will determine the posterior distribution for model parameters. Here response variable also assumed to come from a normal distribution. 

$P(\beta|y,X) = P(y|\beta, X) * P(\beta|X) / P(y|X)$

```{r, eval=FALSE, include=FALSE}
### took too long to run suspended execution Need too see if verbose can be set to true.


train_y_factor = factor(train_y)

model_21_bayes <- BLR(y=train_y, train_x)

summary(model_21_bayes)


#print(paste0("Intercept: ",model_21_bayes$mu))
#print("Coefficents:")
#print(model_21_bayes$bF)



```




###Cubist Model

Cubist is a prediction-oriented regression model that combines the ideas in Quinlan (1992) and Quinlan (1993).  
Although it initially creates a tree structure, it collapses each path through the tree into a rule. A regression model is fit for each rule based on the data subset defined by the rules. The set of rules are pruned or possibly combined. and the candidate variables for the linear regression models are the predictors that were used in the parts of the rule that were pruned away. This part of the algorithm is consistent with the "M5" or Model Tree approach.  
Cubist generalizes this model to add boosting (when committees > 1) and instance based corrections (see predict.cubist()). The number of instances is set at prediction time by the user and is not needed for model building.


```{r, eval=FALSE, cache=TRUE, include=FALSE}

#getting error below

#You are trying to do regression and your outcome only has two possible values Are you trying to do classification? If so, use a 2 level factor as your outcome column.

### prob need to change y to factor

#will suspend from running

set.seed(123)
ctrl=(trainControl(method="repeatedcv", repeats=5))

c<-c(1,3,5,10,20,30,50,80,100)
n<-c(0,1,3,5,7)



model_31_cubit<-train(train_x,train_y, method="cubist",
                preProcess = c("center", "scale"),
                tuneGrid = expand.grid(committees=c,neighbors=n),
                trControl = ctrl)

``` 

Predictor importance from cubist model

```{r, eval=FALSE, cache=TRUE, include=FALSE}

dotPlot(varImp(cube_fit), main="Cubist Predictor importance")

``` 

\begin{center}
fig. 4
\end{center}


### Neural Network Model

In this model, we will try out neural network and perform predictions on using it.

```{r, eval=FALSE, include=FALSE}
#Line 548 erroring out
#due to keras need of python
#will ignore chunk


train_x_s <- scale(train_x)
train_y_cat <- train_y %>% as.matrix() %>% to_categorical()

test_x_s <- scale(test_x)

test_y_cat <- test_y %>% as.matrix()


model_41_neural <- keras_model_sequential() 

model_41_neural %>% layer_dense(units = 8, activation = 'relu', input_shape = c(30)) %>% 
  layer_dense(units = 64, activation = "relu") %>%
  layer_dense(units = 32, activation = "relu") %>%
  layer_dense(units = 2, activation = 'softmax')

model_41_neural %>% compile(loss='categorical_crossentropy',optimizer='rmsprop',metrics='accuracy')

history = model_41_neural %>% fit(train_x_s,train_y_cat, epochs=50,batch_size = 8,validation_split = 0.2)

```

###SVM

I first started by splitting the data and removing predictors with a high correlation. I did this because I could not understand the splitting method but will revise after the group meeting. The pre-processing method was centered and scaled. A CV train control was used. The model yielded an accuracy of 97%. I will return and make adjustments next time.
```{r}


set.seed(143) 
sample = sample.int(n = nrow(WDBCdatafull), size = floor(.70*nrow(WDBCdatafull)), replace = F)
Chris_train = WDBCdatafull[sample, ]
Chris_test  = WDBCdatafull[-sample,] 

corThresh = .9 
tooHigh = findCorrelation(cor(Chris_train), corThresh) 
corrPred = names(Chris_train)[tooHigh]
Chris_train = Chris_train[, -tooHigh]

#removed higly corralated terms from test data set
Chris_test = Chris_test[, -tooHigh]

#Convert diagnosis into factor
Chris_train$diagnosis =as.factor(Chris_train$diagnosis)
Chris_test$diagnosis =as.factor(Chris_test$diagnosis)


set.seed(143)
SVM_Radial_Fit = train(diagnosis~.,Chris_train, method = "svmRadial",
preProc = c("center", "scale"),
tuneLength = 14,trControl = trainControl(method = "cv"))

SVM_Radial_Fit


set.seed(143)
Chris_SVM_Pred = data.frame(Pred= predict(SVM_Radial_Fit, newdata = Chris_test %>% select(-diagnosis)))


plot(SVM_Radial_Fit, scales = list(x = list(log = 2)))

Chris_Con =confusionMatrix(Chris_SVM_Pred$Pred,Chris_test$diagnosis,dnn = c("Prediction", "Reference"))

Chris_SVM_acc = data.frame( Val =postResample(pred = Chris_SVM_Pred$Pred, obs = Chris_test$diagnosis))

fourfoldplot(Chris_Con$table,main=paste("SVM (",round(Chris_SVM_acc[1,]*100),"%)",sep=""),color = c("#ed3b3b", "#0099ff"))

Chris_Con


```




## MODEL ANALYSIS AND DIAGNOSTIC 

In below section, we will test several models and select the best model out of it.


```{r}

# Function for printing confusion matrix
confusion_analysis <- function(df,model){
  # Threshold value is 0.5, positive class is 1
  
  predicted = if_else(predict(model,df)>=0.5, 1,0)
  confusionMatrix(data = predicted,
                reference = df$Diagnosis,
                positive = "1")
}

# Function for calculating evaluation metrics
summary_analysis <- function(df,model){
  print(summary(model))
  print(paste0("BIC: ",BIC(model)))
  print(paste0("VIF: ",vif(model)))

  n = length(df$target)
  print(paste0("Naglekerke-pseudo-R2:",(1-exp((model$dev-model$null)/n))/(1-exp(-model$null/n))))
  print("Confusion Matrix:")
  confusion_analysis(df,model)
}


```


```{r}



confusion_analysis(cbind(test_x,Diagnosis = test_y),model_13_logit_corr_final)
confusion_analysis(cbind(test_x,Diagnosis = test_y),model_15_probit)

#did not run cubit model
#confusion_analysis(cbind(test_x,Diagnosis = test_y),model_31_cubit)


```

```{r include=FALSE, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
#suspended code
rocCurve <- roc(response = test_y,
                predictor = predict(model_31_cubit,newdata = test_x),
                  levels = c(1,0))
```



```{r include=TRUE, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
#will suspend code

print('ROC Curve:')
auc(rocCurve)


plot(y = rocCurve$sensitivities, ylab = "Sensitivity", x = 1 - rocCurve$specificities, xlab = "1 - Specificity",
main = "ROC Curve", col = "red")
```





#Summary and Conclusions

#References

Things to do

#Other

##additional Non-parametric models
##Additional diagnostic plots
##Paperwrite and comments
##Shiny app - if required
##Presentation


REFERENCES:  
[1] Dua, D. and Karra Taniskidou, E. (2017). UCI Machine Learning Repository [http://archive.ics.uci.edu/ml]. Irvine, CA: University of California, School of Information and Computer Science.    
[2] https://stackoverflow.com/questions/7074246/show-correlations-as-an-ordered-list-not-as-a-large-matrix
[3] https://cran.r-project.org/web/packages/Cubist/Cubist.pdf
